Racets: Embedding Policy{-}Agnostic Programming in Racket

### 2018-07-21 paper/paper.tex
Pass over 4

%% To color math, put it in an \mc command, e.g., \mc{$a$}
\newcommand{\colorMATH}{black!20!blue}
\newcommand{\colorFAC}{black!20!red}
\newcommand{\blue}[1] {{\color{\colorMATH} #1}}
\newcommand{\fcol}[1] {{\color{\colorFAC} #1}}
\newcommand{\bmth}[1] {{\color{\colorMATH} $#1$}}
\newcommand{\code}[1]{\lstinline{#1}}

@@

\begin{figure*}
  \hfill \textit{(Metafunctions)}
      \\
      \begin{displaymath}
        \begin{array}{rclrr}
          {read(\sigma, \alpha, pc)} & {=} & {\sigma(\alpha)} & & \\
          {read(\sigma, \facet{\ell}{v_1}{v_2}), pc} & = & {read(\sigma, v_1, pc)} & \textit{if} ~ {+\ell} \in pc & \\
          {read(\sigma, \facet{\ell}{v_1}{v_2}), pc} & = & {read(\sigma, v_2, pc)} & \textit{if} ~ {-\ell} \in pc & \\
          {read(\sigma, \facet{\ell}{v_1}{v_2}), pc} & = & {\formfacet{\ell}{read(\sigma, v_1, pc)}{read(\sigma, v_2, pc)}} & \textit{otherwise} &
        \end{array}
      \end{displaymath}
      \\
      \begin{displaymath}
        \begin{array}{rclrr}
          {write(\sigma, \alpha, pc, v)} & {=} & {\sigma[\alpha ~ := \formfacet{pc}{v}{\sigma(\alpha)}]} & & \\
          {write(\sigma, \facet{\ell}{v_1}{v_2}), pc, v} & = & {\sigma''} & \textit{where} ~ {\sigma' = write(\sigma, v_1, pc \cup \{+\ell\}, v)} \\
          & & & \textit{and} ~ {\sigma'' = write(\sigma', v_2, pc \cup \{-\ell\}, v)}
        \end{array}
      \end{displaymath}
      \\
      \begin{displaymath}
        \begin{array}{rclrr}
          {obs(\ell, b, bv)} & {=} & {bv} & & \\
          {obs(\ell, true,} \facet{\ell}{v_1}{v_2}) & = & {v_1} & & \\
          {obs(\ell, false,} \facet{\ell}{v_1}{v_2}) & = & {v_2} & & \\
          {obs(\ell, b,} \facet{\ell'}{v_1}{v_2}) & = & \facet{\ell'}{obs(\ell,b,v_1)}{obs(\ell,b,v_2)} & \textit{where} & {\ell} \neq {\ell'}
        \end{array}
      \end{displaymath}
\end{figure*}

@@

\begin{figure}
{\small
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets]
(define-syntax-rule (fac l e1 e2)
  (cond
\end{lstlisting}
\vspace{-.7\baselineskip}
{\color{brown}
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets,identifierstyle=\color{brown}]
    [(set-member? (current-pc) (pos (facet-labelname l))) e1]
\end{lstlisting}
}
\vspace{-.7\baselineskip}
{\color{gray}
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets,identifierstyle=\color{gray}]
    [(set-member? (current-pc) (neg (facet-labelname l))) e2]
\end{lstlisting}
}
\vspace{-.7\baselineskip}
{\small\color{purple}
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets,identifierstyle=\color{purple},keywordstyle=\color{purple}]
    [else
     (let ([left (parameterize
                     ([current-pc (set-add (current-pc)
                                           (pos (facet-labelname l)))]))]
           [right (parameterize
                      ([current-pc (set-add (current-pc)
                                            (neg (facet-labelname l)))]))]
           (mkfacet (labelpair-name l) v1 v2)))]))
\end{lstlisting}}
\caption{Facet creation as a macro}
\label{fig:fac-create}
\end{figure}

@@

\paragraph*{Faceted References}

@@

\begin{figure}
(define-syntax (ref-set! stx)
  (syntax-case stx ()
    [(_ var e)
     #`(let ([value e])
         (let setf ([var var]
                    [pc (current-pc)])
           (if (box? var)
               ; if var is a facet value
               (set-box! var (construct-facet-optimized (set->list (current-pc)) value (unbox var)))
               ; else
               (mkfacet
                (facet-labelname (unbox var))
                (setf
                 (facet-left (unbox var))
                 (set-add pc (pos (facet-labelname var))))
                (setf
                 (facet-right (unbox var))
                 (set-add pc (neg (facet-labelname var))))))))]))
\caption{Racets' implementation of \code{set!}}
\label{fig:setbang}
\end{figure}

Racets defines its \code{box} macro, along with \code{unbox} and
\code{set!}. We include the definition of \code{set!} in
Figure~\ref{fig:setbang}, which inlines the definition of the
\textit{write} metafunction from Section~\ref{sec:semantics} to
consider the case under which a facet is used when an address is
expected.

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}

>>>>

%% To color math, put it in an \mc command, e.g., \mc{$a$}
\newcommand{\colorMATH}{black!20!blue}
\newcommand{\colorCOMM}{0.133,0.545,0.133}
\newcommand{\colorFAC}{black!20!red}
\newcommand{\blue}[1] {{\color{\colorMATH} #1}}
\newcommand{\fcol}[1] {{\color{\colorFAC} #1}}
\newcommand{\bmth}[1] {{\color{\colorMATH} $#1$}}
\newcommand{\code}[1]{\lstinline{#1}}

@@

\begin{figure*}
      \begin{displaymath}
        \begin{array}{rclrr}
          {read(\sigma, \alpha, pc)} & {=} & {\sigma(\alpha)} & & \\
          {read(\sigma, \facet{\ell}{v_1}{v_2}), pc} & = & {read(\sigma, v_1, pc)} & \textit{if} ~ {+\ell} \in pc & \\
          {read(\sigma, \facet{\ell}{v_1}{v_2}), pc} & = & {read(\sigma, v_2, pc)} & \textit{if} ~ {-\ell} \in pc & \\
          {read(\sigma, \facet{\ell}{v_1}{v_2}), pc} & = & {\formfacet{\ell}{read(\sigma, v_1, pc)}{read(\sigma, v_2, pc)}} & \textit{otherwise} &
        \end{array}
      \end{displaymath}
      \\
      \begin{displaymath}
        \begin{array}{rclrr}
          {\var{write}(\sigma, \alpha, pc, v)} & {=} & {\sigma[\alpha ~ := \formfacet{pc}{v}{\sigma(\alpha)}]} & & \\
          {\var{write}(\sigma, \facet{\ell}{v_1}{v_2}), pc, v} & = & {\sigma''} & \textit{where} ~ {\sigma' = \var{write}(\sigma, v_1, pc \cup \{+\ell\}, v)} \\
          & & & \textit{and} ~ {\sigma'' = \var{write}(\sigma', v_2, pc \cup \{-\ell\}, v)}
        \end{array}
      \end{displaymath}
      \\
      \begin{displaymath}
        \begin{array}{rclrr}
          {obs(\ell, b, bv)} & {=} & {bv} & & \\
          {obs(\ell, true,} \facet{\ell}{v_1}{v_2}) & = & {v_1} & & \\
          {obs(\ell, false,} \facet{\ell}{v_1}{v_2}) & = & {v_2} & & \\
          {obs(\ell, b,} \facet{\ell'}{v_1}{v_2}) & = & \facet{\ell'}{obs(\ell,b,v_1)}{obs(\ell,b,v_2)} & \textit{where} & {\ell} \neq {\ell'}
        \end{array}
      \end{displaymath}
      \caption{Meta-functions used in our semantics}
      \label{fig:metafunctions}
\end{figure*}

@@

\begin{figure}
{\small
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets]
(define-syntax-rule (fac l e1 e2)
  (cond
\end{lstlisting}
\vspace{-.7\baselineskip}
{\color{brown}
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets,identifierstyle=\color{brown}]
    [(set-member? (current-pc) (pos (facet-labelname l))) e1]
\end{lstlisting}
}
\vspace{-.7\baselineskip}
{\color{gray}
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets,identifierstyle=\color{gray}]
    [(set-member? (current-pc) (neg (facet-labelname l))) e2]
\end{lstlisting}
}
\vspace{-.7\baselineskip}
{\small\color{purple}
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets,identifierstyle=\color{purple},keywordstyle=\color{purple}]
    [else
     (let ([left (parameterize
                     ([current-pc (set-add (current-pc)
                                           (pos (facet-labelname l)))]))]
           [right (parameterize
                      ([current-pc (set-add (current-pc)
                                            (neg (facet-labelname l)))]))]
           (mkfacet (labelpair-name l) v1 v2)))]))
\end{lstlisting}}
}
\caption{Facet creation as a macro}
\label{fig:fac-create}
\end{figure}

@@

\paragraph*{Faceted Boxes, Writes, and Observations}

@@

\begin{figure}
\small
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets]
(define-syntax (ref-set! stx)
  (syntax-case stx ()
    [(_ var e)
     #`(let ([v e])
         (let write ([var var]
                     [pc (current-pc)])
           (if (box? var)
               ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc, v)$@>
               (set-box! var (construct-facet (current-pc) v (unbox var)))
               ; Else split
               (mkfacet
                 (facet-labelname (unbox var))
                 ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc \cup \{+l\}, v)$@>
                 (write
                   (facet-left (unbox var))
                   (set-add pc (pos (facet-labelname var))))
                 ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc \cup \{-l\}, v)$@>
                 (write
                   (facet-right (unbox var))
                   (set-add pc (neg (facet-labelname var))))))))]))
\end{lstlisting}
\caption{Racets' implementation of \code{set!}}
\label{fig:setbang}
\end{figure}

Racets defines a \code{box} macro, along with \code{unbox} and
\code{set!}. We include the definition of \code{set!} in
Figure~\ref{fig:setbang}, which inlines the definition of the
\textit{write} metafunction from Section~\ref{sec:semantics} to
consider the case under which a facet is used when an address is
expected.

Facet observation is handled similarly, first evaluating the label to
produce a policy predicate, followed by evaluating the policy's
argument and a possibly-faceted value to observe. After applying the
policy its argument, we produce the value \bmth{v^{\pm}} and descend
down the facet until reaching either a base value or finding the
selected label (at which point we select the appropriate branch).

\paragraph*{Faceted Function Application}

By now we can anticipate a predictable pattern for implementing
faceted execution: check \bmth{pc} to decide whether to branch left,
right, or split. This is largely our strategy for handling faceted
function application, with a small twist: we need to be able to apply
functions from outside of Racets. For example, if we want to apply
builtin functions such as \code{display}, we need to be mindful of the
fact that these functions cannot work with faceted arguments.

To handle this, we implement a macro for the \code{lambda} form, to
tag Racet closures specifically (so that they are differentiated from
functions outside of the current module). In the case that a foreign
function is applied to a faceted value, our implementation of function
application wraps the function to be able to handle facets by
distributing the function through each branch of the facet.

In general it is unsafe to apply an unknown function to a faceted
value. This is because the unknown function may leak the facet's
private information as a side-effect. Therefore, our current
implementation of Racets allows programmers to apply external
functions, but does not make any guarantee of safety. A better
strategy may be to perform an \code{obs} before each call to a
potentially-unsafe function. In general, we believe module
interactions are a challenging problem in faceted execution, and we
leave its study to future work.

\section{Implementation and Evaluation}

\section{Conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}
\end{document}
